---
title: "Assignment 1"
author: "Lin Ye"
date: "January 7, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


####1. Write a loop that calculates 12-factorial 

First, get user input. Enter -12.1 as an example



```{r}
num_entry = as.numeric(readline(prompt="Enter a number: "))

num_entry =-12.1

##check if the number is negative or decimal

if(num_entry!=round(num_entry) || num_entry < 0){
  print("Factorial cannot be non integer or negative.")
  num = abs(round(num_entry,0) )
} else {
    num=num_entry  
}
```

Once the entry passed checks, print the new number

```{r}
print(paste("You entered",num_entry,". Showing result for", num))
factorial = 1
```

Last, calculate factorial 

```{r}
if(num == 0) {
  print("The factorial of 0 is 1")
} else {
  for(i in 1:num) {
    factorial = factorial * i
  }
  print(paste("The factorial of", num ,"is",factorial))
}
``` 


####2. Show how to create a numeric vector that contains the sequence from 20 to 50 by 5. 

```{r}
x = 20
y = 50
z = 5
seq(x, y, z)
```

####3. Create the function "factorial" that takes a trio of input numbers a, b, and c and solve the quadratic equation.  The function should print as output the two solutions. 

User define funations

```{r}

factorial_x1 <- function(a,b,c){
  (-b+sqrt(m))/2*a
}

factorial_x2 <- function(a,b,c){
  (-b-sqrt(m))/2*a
}


```


Checks are made to ensure a > 0 and no imaginary numbers in calculation.

First, we test it with a=0, b=2, c=1

```{r}
a=0
b=2
c=1
m=(b*b)-4*a*c

if(a != 0 && m > 0){
  x1 <-factorial_x1(a,b,c)
  x2 <-factorial_x2(a,b,c)
  factorial<-data.frame(cbind(a,b,c,"Solution1"=x1, "Solution2"=x2))
  print(factorial)
}else{
  print("Yields invalid result, try again...")
}
```

Next, we test it with a=2, b=4, c=1

```{r}
a=2
b=4
c=1
m = (b*b)-4*a*c

if(a != 0 && m > 0){
  x1 <-factorial_x1(a,b,c)
  x2 <-factorial_x2(a,b,c)
  factorial<-data.frame(cbind(a,b,c,"Solution1"=x1, "Solution2"=x2))
  print(factorial)
}else{
  print("Yields invalid result, try again...")
}
```
