---
title: "Winter 2017 Week 2"
author: "Lin Ye"
date: "January 15, 2017"
output: html_document
---

In this assignment, I'm looking at the rating data for US judges

####Description

##Lawyers' ratings of state judges in the US Superior Court. 
Document is found here http://vincentarelbundock.github.io/Rdatasets/doc/datasets/USJudgeRatings.html

Two ways to get the data
1. Through file on local drive
```{r}
setwd("C:/Users/liye/Documents/R Class")
getwd()
data1= read.csv("USJudgeRatings.csv")

head(data1)

rm(data1)
```

2.Read from link on github.
```{r}
data <- read.csv(url("http://vincentarelbundock.github.io/Rdatasets/csv/datasets/USJudgeRatings.csv"))

#Remove some columns we don't need and only focus on a few attributes
data_new <-subset(data[1:11],select=-c(6:8))

print(paste(length(data_new), "columns and",nrow(data_new),"observations are found in the judge rating dataset"))


#Renaming fields to corresponding names based on dictionary
colnames(data_new) <- c("Name","NumOfContacts","Integrity","Demeanor","Diligence","Familiarity","Oral","Written")

#summerize first three attributes of the dataset
summary(data_new[2:7])

#Taking a subset of data where # times of contact is above 7
sub_data <-data_new[data_new[,3]>=8,]

print(paste(length(sub_data), "columns and",nrow(sub_data),"observations are found in the judge rating dataset"))

summary(sub_data[2:7])


#mean of the attributes and remove missing values
sapply(data_new[4:8], mean, na.rm=TRUE)

sapply(sub_data[4:8], mean, na.rm=TRUE)

#median of the attributes and remove missing values
sapply(data_new[4:8], median, na.rm=TRUE) 

sapply(sub_data[4:8], median, na.rm=TRUE) 

```

Based on the summaries above, judges that have higher integrity are on average receiving higher rating across all other areas. The median of all the measures is also higher among judges with higher integrity.

```{r}

data_new["Integrity_Ranking"]<-data_new$Integrity

data_new$Integrity_Ranking <- with(data_new, ifelse(Integrity < 7, "Low",
                     ifelse(Integrity >=7 & Integrity < 8, "Med",
                      "High")))

data_new_1 <-data_new[,c(1:3,9,4:8)]

head(data_new_1,20)
```

